Case study: Analysis of smart meter data

### 2019-08-26 smart_meter.Rmd
smart meter analysis uploaded

Figure \ref{fig:cust1_hd_my} and \ref{fig:cust3_hd_my} shows the distribution of hourly consumption of household 1 and 2 across different months of the year. The two households are similar when it comes to their usage in the morning hours. The first household's extreme behavior is different from that of the second household in the evening hours of May and June (represented by higher 90th percentile energy usage). However, the 75th percentile of energy consumption for those two months are higher for the second household than for the first and their median consumption is more or less similar. This implies that consuming more energy in the evening is more typical for household 2. For the evenings of October, November, and December, it is clear that household 1 typically consumes much less energy than household 2. Moreover, the second household has a steep evening peak (75th percentile above 0.50 KwH) across most months in contrast to the first household who has a steep evening peak only in the month of May and June with 75th percentile consumption falling below 0.25 KwH. This might be an indication that of usage of a heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely. All of these can be explained with more demographics of the household and weather conditions in that area.

@@

```{r cust1_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the first household. 10th, 25th, 50th, 75th and 90th percentile are shown. For example, 90th percentile might denote behavior which are extreme and 50th percentile corresponds to a more typical behavior. Evening hours generally exhibits a peak. Some extreme behavior is exhibited in the evening hours of May and June, however, overall energy comsumption does not seem to vary much across months. Median and 75th percentile is less than 0.25 KwH across most months."}
library(ggplot2)
data_sm1 %>% 
  granplot("month_year", "hour_day", 
           response = "general_supply_kwh", 
           plot_type = "quantile", 
           quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  #ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +ggtitle("") +
  scale_x_discrete(breaks = seq(0,23,2))
```

```{r cust3_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the second household. 10th, 25th, 50th, 75th and 90th percentile are shown. Distinct evening peaks for few hours across all months. The median comsumption also varies across months. This household varies their behavior with seasons and are likely to be using heaters in colder months and air conditioners in hot months like January. This is evident from the 75th and 90th percentile of consumption."}
data_sm3 %>% 
  create_gran("month_year") %>% 
  filter(month_year != "Mar") %>% 
    granplot("month_year", "hour_day", "general_supply_kwh",
             plot_type = "quantile", 
             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("") +
  #ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 2") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
  scale_x_discrete(breaks = seq(0,23,2))
```

@@

Figure \ref{fig:cust1_hd_dw} shows the distribution of hourly energy usage across different days of the week through violin plots for household 1. Energy consumption typically goes up between 16 to 18 hours. On weekends, the energy consumption for these hours is even higher. The straight lines in the violin plot for weekends at 17th and 18th hours suggest that the distribution of energy consumption is uniform demonstrating that any energy value has equal probability in falling in that hour. This suggests that these hours are extremely volatile based on weather conditions/months. Energy consumption distribution in the morning hours of weekdays has a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical weekday, we can say with reasonable certainty that the energy consumption would fall in that range. This is not true for the 17th or 18th hour of the day. The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home parents/ working from home professionals.

Figure \ref{fig:cust3_hd_dw} shows the distribution of hourly energy usage across different days of the week through violin plots for Household 2. Contrary to household 1, household 2's energy consumption increases through noon to 7 pm. We can't say with certainty if the morning hours from 10 am would manifest a higher or lower level of energy consumption for this household. This indicates that this household might have kids or someone working from home.

```{r cust1_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the first household.  A long-tailed violin for an hour would imply that the energy distribution is skewed to either right or left, whereas a bulge shaped distribution implies that the energy behavior is more predictable and energy consumption in those hours can only vary in that range where the bulge appears. If an hour has a distribution that seems like a straight line, they are extremely volatile. On weekdays, hours between 17 to 19 hours seems to be extremely volatile compared to other hours. On weekends, energy consumption is volatile even in morning hours. "}

data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("") +
  #ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") 
  scale_x_discrete(breaks = seq(0,23,3)) +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```
<!-- # ``````{r cust2_hour-day_day-week} -->
<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
<!-- # ``` -->

```{r cust3_hd_dw, fig.cap ="Violin plots of hourly consumption across different day of the week for the second household. Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent. Hours in between these have a long right tail which are even higher for weekends."}
data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +
  #ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") +
  scale_x_discrete(breaks = seq(0,23,3)) +
  ggtitle("") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

@@

```{r cust1_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household. Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations. M, F, E, D and C represents 50, 25, 12.5, 6.25 and 3.13 percent of the tail area respectively.Outliers are colored in red. It can be noted that they do not typically consume more than 0.25 KwH in the weekdays. Their tail behaviors are not so distinctly visible because of the consistency in their behavior."}

data_sm1 %>% granplot("month_year", 
                      "day_week", 
                      response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +  #ggtitle("LV plot of energy consumption across days-week by month-year for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

```{r cust3_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household. Contrary to the first household, outliers appear at an energy usage beyond 0.75 KwH. The extreme values in the tail of the distribution are represented by different letter values. In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays -represented by letter value D, which is far more higher than subsequent months and comparable to summer month December. For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."}

data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
  #ggtitle("LV plot of energy consumption across days of week by month of years for Household 2") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

@@

Figure \ref{fig:cust1_wm_wndwd} shows that median consumption of energy is higher on Sundays. Also Sundays have longer right-tails than Saturdays across all weeks of the month implying that Sunday is a little more active day for the household. And the forth weekend of the month, either they take it easy on themselves or stay out of their house for trips/visit friends.

Figure \ref{fig:cust3_wm_wndwd} shows that household 2 has all-absorbing Sundays. They are either busy with weekly chores or have the entire family stay at home every Sunday. The first Saturday of the month looks pretty consistent compared to any other weeks of the month, implying either they are out for work/trips/visiting family. Also, 4th Sunday of the month has a long right tail implying they choose the last week of the month for doing their "monthly" chores.

```{r cust1_wm_wndwd, fig.cap = "Letter value plot of energy consumption for Saturday and Sunday across weeks of the month for household 1. The 4th week of the month sees low energy consumption for both Saturdays and Sundays.Typically Sundays are heavier on energy than Saturdays."}

data_sm1 %>% 
   create_gran("day_week") %>% 
  filter(day_week %in%c("Sat", "Sun")) %>% 
  granplot("day_week", "week_month" , response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
  #ggtitle("Boxplot plot of energy consumption across days of week by month of years for Household 2") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

```{r cust3_wm_wndwd, fig.cap = "Letter value plot of energy consumption for Saturday and Sunday across weeks of the month for household 2. The forth  Sunday has high energy consumption and long right tails implying they choose the last Sunday of the month either to do monthly chores or have friends come over. The first Saturday of the week seems quite and consistent for this household."}

data_sm3 %>%
  create_gran("day_week") %>% 
  filter(day_week %in%c("Sat", "Sun")) %>% 
  granplot("day_week", "week_month" , response = "general_supply_kwh", plot_type = "lv") + ggtitle("") +
  #ggtitle("Boxplot plot of energy consumption across days of week by month of years for Household 2") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

@@

We can keep on slicing and dicing the data similarly and derive more insights about the behavior of these households.

>>>>

Figure \ref{fig:cust1_hd_my} shows the distribution of hourly consumption of both households across different months of the year. The two households are similar when it comes to their usage in the morning hours. The first household's extreme behavior is different from that of the second household in the evening hours of May and June (represented by higher 90th percentile energy usage). However, the 75th percentile of energy consumption for those two months are higher for the second household than for the first and their median consumption is more or less similar. This implies that consuming more energy in the evening is more typical for household 2. For the evenings of October, November, and December, it is clear that household 1 typically consumes much less energy than household 2. Moreover, the second household has a steep evening peak (75th percentile above 0.50 KwH) across most months in contrast to the first household who has a steep evening peak only in the month of May and June with 75th percentile consumption falling below 0.25 KwH. This might be an indication that of usage of a heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely. All of these can be explained with more demographics of the household and weather conditions in that area.

@@

```{r cust1_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the first household. 10th, 25th, 50th, 75th and 90th percentile are shown. For example, 90th percentile might denote behavior which are extreme and 50th percentile corresponds to a more typical behavior. Some extreme behavior is exhibited in the evening hours of May and June for household 1, however, overall energy comsumption does not seem to vary much across months. Median and 75th percentile is less than 0.25 KwH across most months. For household 2, distinct evening peaks for observed for few evening hours across all months. The median comsumption also varies across months. They are likely to be using heaters in colder months and air conditioners in hot months like January."}
library(ggplot2)
data_sm1 %>% 
  granplot("month_year", "hour_day", 
           response = "general_supply_kwh", 
           plot_type = "quantile", 
           quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  ggtitle("Quantile plot across hour-of-day by month-ofyear for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +ggtitle("") +
  scale_x_discrete(breaks = seq(0,23,2))

data_sm3 %>% 
  create_gran("month_year") %>% 
  filter(month_year != "Mar") %>% 
    granplot("month_year", "hour_day", "general_supply_kwh",
             plot_type = "quantile", 
             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  +
  ggtitle("Quantile plot across hour-of-day by month-ofyear for Household 1") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
  scale_x_discrete(breaks = seq(0,23,2))
```

<!-- # ```{r cust3_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the second household. 10th, 25th, 50th, 75th and 90th percentile are shown. Distinct evening peaks for few hours across all months. The median comsumption also varies across months. This household varies their behavior with seasons and are likely to be using heaters in colder months and air conditioners in hot months like January. This is evident from the 75th and 90th percentile of consumption."} -->
<!-- #  -->
<!-- # ```  -->

@@

Figure \ref{fig:cust1_hd_dw} shows the distribution of hourly energy usage across different days of the week through violin plots for both households. Energy consumption typically goes up between 16 to 18 hours. On weekends, the energy consumption for these hours is even higher. The straight lines in the violin plot for weekends at 17th and 18th hours suggest that the distribution of energy consumption is uniform demonstrating that any energy value has equal probability in falling in that hour. This suggests that these hours are extremely volatile based on weather conditions/months. Energy consumption distribution in the morning hours of weekdays has a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical weekday, we can say with reasonable certainty that the energy consumption would fall in that range. This is not true for the 17th or 18th hour of the day. The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home parents/ working from home professionals.

For the second household, the distribution of hourly energy usage across different days of the week through violin plots for Household 2. Contrary to household 1, household 2's energy consumption increases through noon to 7 pm. We can't say with certainty if the morning hours from 10 am would manifest a higher or lower level of energy consumption for this household. This indicates that this household might have kids or someone working from home.

```{r cust1_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the two households  A long-tailed violin for an hour would imply that the energy distribution is skewed to either right or left, whereas a bulge shaped distribution implies that the energy behavior is more predictable and energy consumption in those hours can only vary in that range where the bulge appears. If an hour has a distribution that seems like a straight line, they are extremely volatile. For household 1, hours between 17 to 19 hours (weekdays) seem to be volatile compared to other hours. On weekends, energy consumption is volatile even in morning hours. For the second one, hours before 10am and beyond 10pm have a bulge implying those hours are pretty consistent. Hours in between these are variable."}

data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("") +
  ggtitle("Violin plot across hour-of-day and day-of-week for Household 1") +
  scale_x_discrete(breaks = seq(0,23,3)) +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))

data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +
  ggtitle("Violin plot across hour-of-day and day-of-week for Household 2") +
  scale_x_discrete(breaks = seq(0,23,3)) +
  ggtitle("") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```
<!-- # ``````{r cust2_hour-day_day-week} -->
<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
<!-- # ``` -->

<!-- # ```{r cust3_hd_dw, fig.cap ="Violin plots of hourly consumption across different day of the week for the second household. Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent. Hours in between these have a long right tail which are even higher for weekends."} -->
<!-- #  -->
<!-- # ``` -->

@@

```{r cust1_my_dw, fig.cap = "LV plots across days of the week and months. They convey detailed information about tails where outliers (colored in red) are unexpected observations rather than extreme ones. M, F, E, D and C represents 50, 25, 12.5, 6.25 and 3.13 percent of the tail area respectively. Household 1's tail behaviors are not distinctly visible because of the consistency in behavior on weekdays. For the second household, the F letter value seem to range between 0.75 to 1.25 KwH on weekends and 0.25 to 0.75 KwH on weekdays in Mar, which is very different from Dec. This reinforces the fact that their behavior changes across months but not so much between days-of-week."}

data_sm1 %>% granplot("month_year", 
                      "day_week", 
                      response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +  
  ggtitle("LV plot across day-of-week by month-of-year for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))

data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
  ggtitle("LV plot across day-of-week by month-of-year for Household 2")  + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

<!-- # ```{r cust3_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household. Contrary to the first household, outliers appear at an energy usage beyond 0.75 KwH. The extreme values in the tail of the distribution are represented by different letter values. In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays -represented by letter value D, which is far more higher than subsequent months and comparable to summer month December. For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."} -->
<!-- #  -->
<!-- #  -->
<!-- # ``` -->

@@

Figure \ref{fig:cust1_wm_wndwd} shows that median consumption of energy for both households. For household 1,  it is higher on Sundays. Also Sundays have longer right-tails than Saturdays across all weeks of the month implying that Sunday is a little more active day for the household. And the forth weekend of the month, either they take it easy on themselves or stay out of their house for trips/visit friends. Household 2 has all-absorbing Sundays. They are either busy with weekly chores or have the entire family stay at home every Sunday. The first Saturday of the month looks pretty consistent compared to any other weeks of the month, implying either they are out for work/trips/visiting family. Also, 4th Sunday of the month has a long right tail implying they choose the last week of the month for doing their "monthly" chores.

```{r cust1_wm_wndwd, fig.cap = "LV plot for Saturday and Sunday across weeks-of-month. For household 1, the 4th week of the month sees low energy consumption for both days. Typically Sundays are heavier on energy than Saturdays. For household 2, the forth  Sunday has high energy consumption and long right tails implying more variable behavior. The first Saturday of the week seems quiet and consistent for this household."}

data_sm1 %>% 
   create_gran("day_week") %>% 
  filter(day_week %in%c("Sat", "Sun")) %>% 
  granplot("day_week", "week_month" , response = "general_supply_kwh", plot_type = "lv") +  ggtitle("") +
  ggtitle("Letter value plot across day-of-week by week-of-month for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))

data_sm3 %>%
  create_gran("day_week") %>% 
  filter(day_week %in%c("Sat", "Sun")) %>% 
  granplot("day_week", "week_month" , response = "general_supply_kwh", plot_type = "lv") + ggtitle("") +
    ggtitle("Letter value plot across day-of-week by week-of-month for Household 2") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

<!-- # ```{r cust3_wm_wndwd, fig.cap = "Letter value plot of energy consumption for Saturday and Sunday across weeks of the month for household 2. The forth  Sunday has high energy consumption and long right tails implying they choose the last Sunday of the month either to do monthly chores or have friends come over. The first Saturday of the week seems quite and consistent for this household."} -->
<!-- # -->
<!-- # -->
<!-- # ``` -->
<!-- # -->

@@
